Expression QTL hotspot dissection with mediation analysis and pleiotropy testing

### 2019-07-23 analysis/paper/figures.Rmd
updated ms

```{r, scatter_by_local-2}
p <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) %>%
  ggplot() + 
  geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) + 
  facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) +
  broman::karl_theme(strip.placement = "inside") + 
  theme(legend.position = "none") +
  xlab("Pleiotropy LOD") + 
  ylab("Mediation LOD difference")
fnsvg <- paste0("Chr", hot_chr, "_scatter-panel.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
ggsave(plot = p, filename = file.path("../figures", fnsvg), height = ceiling(nlocal / 2))
ggsave(plot = p, filename = file.path("../figures", fnpdf), height = ceiling(nlocal / 2))
```

>>>>

```{r, scatter_by_local-2}
foo <- pleio_mixed_list %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  inner_join(med_tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  dplyr::select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id)
p <- foo %>%
  ggplot() + 
  geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) + 
  facet_wrap(. ~ local_symbol, nrow = ceiling(nlocal / 2), ncol = 2) +
  broman::karl_theme(strip.placement = "inside") + 
  theme(legend.position = "none") +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()
) + 
  xlab("Pleiotropy LOD") + 
  ylab("Mediation LOD difference") + 
  geom_text(aes(x = 1, y = 1, label = local_symbol), inherit.aes=FALSE)

fnsvg <- paste0("Chr", hot_chr, "_scatter-panel.svg")
fnpdf <- paste0("Chr", hot_chr, "_scatter-panel.pdf")
ggsave(plot = p, filename = file.path("../figures", fnsvg), height = ceiling(nlocal / 2))
ggsave(plot = p, filename = file.path("../figures", fnpdf), height = ceiling(nlocal / 2))

# put in multiple figures, if there are many local traits
n_figs <- 1 + nlocal %/% 18 # 18 is the number of panels per page in pdf
n_last <- nlocal %% 18 # number of panels on last figure
max_pleio_lod <- max(foo$pleiotropy_lod)
max_med_lod_difference <- max(foo$lod_diff) 
min_med_lod_difference <- min(foo$lod_diff)

for (i in 1:n_figs){
  n_panels <- 18 * (i < n_figs) + n_last * (i == n_figs)
  foo %>%
    filter(local_symbol %in% local_annot2$symbol[((i - 1) * 18 + 1): (18 * i)]) %>%
    ggplot() + 
      geom_point(mapping = aes(x = pleiotropy_lod, y = lod_diff, colour = nonlocal_symbol, nonlocal = nonlocal_symbol), size = 0.1) + 
      facet_wrap(. ~ local_symbol, nrow = ceiling(n_panels / 2), ncol = 2) +
      broman::karl_theme() + 
      theme(legend.position = "none") +
      theme(
        strip.background = element_blank(),
        strip.text.x = element_blank()
    ) + 
      xlab("Pleiotropy LOD") + xlim(c(0, max_pleio_lod)) + ylim(c(min_med_lod_difference, max_med_lod_difference)) + 
      ylab("Mediation LOD difference") + 
      geom_text(vjust = 0.9, hjust = 0.9, size = 3, aes(x = max_pleio_lod / 2, y = 2 * max_med_lod_difference / 3, label = local_symbol), inherit.aes=FALSE)
  fnsvg <- paste0("Chr", hot_chr, "_scatter-panel", i, ".svg")
  fnpdf <- paste0("Chr", hot_chr, "_scatter-panel", i, ".pdf")
  ggsave(filename = file.path("../figures", fnsvg), height = 9)
  ggsave(filename = file.path("../figures", fnpdf), height = 9)
}
```

### 2019-07-23 analysis/paper/paper.Rnw
updated ms

<<>>=
include_graphics("../figures/Chr2_scatter.pdf")
@

>>>>

<<>>=
hot_chr <- 2
@

<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 2 hotspot.">>=
include_graphics(paste0("../figures/Chr", hot_chr, "_scatter.pdf"))
@

<<define_nfig>>=
nfig <- c(2, 2, 3, 3, 1) # number of panel figs per hotspot
names(nfig) <- paste0("Chr", c(2, 5, 7, 11, 13))
@

<<panel1, fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")>>=
include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel1.pdf"))
@

<<panel2, eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1>>=
include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel2.pdf"))
@

<<panel3, eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2>>=
include_graphics(paste0("../figures/Chr", hot_chr, "_scatter-panel3.pdf"))
@

<<>>=
hot_chr <- 5
@

<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 5 hotspot.">>=
include_graphics("../figures/Chr5_scatter.pdf")
@

<<ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")>>=
@

<<ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1>>=
@

<<ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2>>=
@

<<>>=
hot_chr <- 7
@

<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 7 hotspot.">>=
include_graphics("../figures/Chr7_scatter.pdf")
@

<<ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")>>=
@

<<ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1>>=
@

<<ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2>>=
@

<<>>=
hot_chr <- 11
@

<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 11 hotspot.">>=
include_graphics("../figures/Chr11_scatter.pdf")
@

<<ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")>>=
@

<<ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1>>=
@

<<ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2>>=
@

<<>>=
hot_chr <- 13
@

<<fig.cap="Scatter plot of mediation LOD difference against pleiotropy LOD for all nonlocal-local trait pairs mapping to Chromosome 13 hotspot.">>=
include_graphics("../figures/Chr13_scatter.pdf")
@

<<ref.label = "panel1", fig.cap = paste0("Chr", hot_chr, "hotspot scatter plots of mediation LOD difference against pleiotropy LOD, separated by local trait.")>>=
@

<<ref.label = "panel2", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 1>>=
@

<<ref.label = "panel3", eval = nfig[which(names(nfig) == paste0("Chr", hot_chr))] > 2>>=
@
